---
title: "Validate the bias correction part of DCATS"
author: "Xinyi_Lin"
date: "07/11/2022"
output: html_document
---

## Ritchie 2019

```{r}
library(scater)
library(Seurat)
```

This data comes from 'Benchmarking single cell RNA-sequencing analysis pipelines using mixture control experiments' (https://www.nature.com/articles/s41592-019-0425-8#data-availability). It contains three/five cell lines' cells with same proportion.

```{r}
load("/storage/holab/linxy/DCATS/real_world/Ritchie_2019/9cellmix_qc.RData")
load("/storage/holab/linxy/DCATS/real_world/Ritchie_2019/sincell_with_class_5cl.RData")
```

This is the pseudo cell. Can't find its use right now

```{r, eval=FALSE}
sce_SC1_qc = logNormCounts(sce_SC1_qc)
pseudocell_mix1 = as.Seurat(sce_SC1_qc)
pseudocell_mix1 <- NormalizeData(pseudocell_mix1)
pseudocell_mix1 <- FindVariableFeatures(pseudocell_mix1, selection.method = "vst", nfeatures = 2000)
pseudocell_mix1 <- ScaleData(pseudocell_mix1, verbose = FALSE)
pseudocell_mix1 <- RunPCA(pseudocell_mix1, npcs = 30, verbose = FALSE)
pseudocell_mix1 <- RunUMAP(pseudocell_mix1, reduction = "pca", dims = 1:30)
pseudocell_mix1 <- FindNeighbors(pseudocell_mix1, reduction = "pca", dims = 1:30)
pseudocell_mix1 <- FindClusters(pseudocell_mix1)
DimPlot(pseudocell_mix1)
```

This is the cell pool mixing five cell lines with equal proportion.

```{r}
mixcell5_10x_seurat = as.Seurat(sce_sc_10x_5cl_qc)
mixcell5_10x_seurat <- NormalizeData(mixcell5_10x_seurat)
mixcell5_10x_seurat <- FindVariableFeatures(mixcell5_10x_seurat, selection.method = "vst", nfeatures = 2000)
mixcell5_10x_seurat <- ScaleData(mixcell5_10x_seurat, verbose = FALSE)
mixcell5_10x_seurat <- RunPCA(mixcell5_10x_seurat, npcs = 30, verbose = FALSE)
mixcell5_10x_seurat <- RunUMAP(mixcell5_10x_seurat, reduction = "pca", dims = 1:30)
mixcell5_10x_seurat <- FindNeighbors(mixcell5_10x_seurat, reduction = "pca", dims = 1:30)
DimPlot(mixcell5_10x_seurat, group.by = 'cell_line_demuxlet')
save(mixcell5_10x_seurat, file = "/storage/holab/linxy/DCATS/real_world/Ritchie_2019/mixcell5_10x_seurat.RData")
```

```{r, eval=FALSE}
sc_Celseq2_5cl_p1 = logNormCounts(sc_Celseq2_5cl_p1)
mixcell5_p1_seurat = as.Seurat(sc_Celseq2_5cl_p1)
mixcell5_p1_seurat <- NormalizeData(mixcell5_p1_seurat)
mixcell5_p1_seurat <- FindVariableFeatures(mixcell5_p1_seurat, selection.method = "vst", nfeatures = 2000)
mixcell5_p1_seurat <- ScaleData(mixcell5_p1_seurat, verbose = FALSE)
mixcell5_p1_seurat <- RunPCA(mixcell5_p1_seurat, npcs = 30, verbose = FALSE)
mixcell5_p1_seurat <- RunUMAP(mixcell5_p1_seurat, reduction = "pca", dims = 1:30)
mixcell5_p1_seurat <- FindNeighbors(mixcell5_p1_seurat, reduction = "pca", dims = 1:30)
mixcell5_p1_seurat <- FindClusters(mixcell5_p1_seurat)
DimPlot(mixcell5_p1_seurat, group.by = 'cell_line_demuxlet')
```

Try to merge p1-p3 together. Check whether they have batch effect.

Results: have batch between technologies. Since 10x has more cell numbers, use this one

```{r, eval=FALSE}
sc_Celseq2_5cl_p2 = logNormCounts(sc_Celseq2_5cl_p2)
mixcell5_p2_seurat = as.Seurat(sc_Celseq2_5cl_p2)
sc_Celseq2_5cl_p3 = logNormCounts(sc_Celseq2_5cl_p3)
mixcell5_p3_seurat = as.Seurat(sc_Celseq2_5cl_p3)
```

```{r, eval=FALSE}
mixcell5_merge = merge(mixcell5_p1_seurat, c(mixcell5_p2_seurat, mixcell5_p3_seurat, mixcell5_10x_seurat), add.cell.ids = c("p1", "p2", "p3", "10x"), project = "mixcell5")
mixcell5_merge <- NormalizeData(mixcell5_merge)
mixcell5_merge <- FindVariableFeatures(mixcell5_merge, selection.method = "vst", nfeatures = 2000)
mixcell5_merge <- ScaleData(mixcell5_merge, verbose = FALSE)
mixcell5_merge <- RunPCA(mixcell5_merge, npcs = 30, verbose = FALSE)
mixcell5_merge <- RunUMAP(mixcell5_merge, reduction = "pca", dims = 1:30)
mixcell5_merge <- FindNeighbors(mixcell5_merge, reduction = "pca", dims = 1:30)
DimPlot(mixcell5_merge, group.by = 'cell_line_demuxlet')
DimPlot(mixcell5_merge, group.by = 'batch')
```

