---
title: "Simulation_plus5_2"
output: html_document
---

Try theoretical simulation for figure 1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
```

```{r,message=FALSE,warning=FALSE}
library(splatter)
library(Seurat)
library(speckle)
library(DCATS)
library(ggplot2)
library(tidyverse)
library(MCMCpack)
library(pROC)
library(confusionMatrix)
```

```{r}
source("functionsV2.r")
```

## Figure draft part

Fig 1 (theoretical simulation might be enough): the usefulness of using similarity matrix to correct the bias;

Draw a boxplot to show AUC results of DCATS, chi-square, scDC(, fisher) without BC, with BC and with BC using resampling.

```{r}
cluster_num = 7  # numbers of clusters
concentration = 50 # indicate how simulated proportion far away from true, the larger the closer
rep1 = 6
rep2 = 4
simulation_times = 50
simulation_size = 50
sample_size1 = 1000
sample_size2 = 2000
```

```{r}
if (cluster_num == 3){
  probC1 = c(1/3, 1/3, 1/3)  # True proportions
  probC2 = c(1/3, 1/2, 1/6)
  truthRes = c("N", "P", "P")
} else if (cluster_num == 7){
  probC1 = c(rep(0.1, 4), rep(0.2, 3))
  probC2 = c(0.1, 0.1, 0.2, 0.05, 0.2, 0.3, 0.05)
  truthRes = c("N", "N", "P", "P", "N", "P", "P")
} else if (cluster_num == 12) {
  probC1 = c(rep(0.1, 4), rep(0.2, 2), rep(0.05, 2), rep(0.025, 4))
  probC2 = c(0.1, 0.1, 0.15, 0.05, 0.2, 0.1, 0.05, 0.1, 0.025, 0.025, 0.05, 0.05)
  truthRes = c("N", "N", "P", "P", "N", "P", "N", "P", "N", "N", "P", "P")
}
```

```{r}
## more negative
if (cluster_num == 3){
  probC1 = c(1/3, 1/3, 1/3)  # True proportions
  probC2 = c(1/3, 1/2, 1/6)
  truthRes = c("N", "P", "P")
} else if (cluster_num == 7){
  probC1 = c(rep(0.1, 4), rep(0.2, 3))
  probC2 = c(0.1, 0.1, 0.1, 0.15, 0.2, 0.3, 0.05)
  truthRes = c("N", "N", "N", "P", "N", "P", "P")
} else if (cluster_num == 12) {
  probC1 = c(rep(0.1, 4), rep(0.2, 2), rep(0.05, 2), rep(0.025, 4))
  probC2 = c(0.1, 0.1, 0.1, 0.1, 0.2, 0.15, 0.05, 0.075, 0.05, 0.025, 0.025, 0.025)
  truthRes = c("N", "N", "N", "N", "N", "P", "N", "P", "P", "N", "N", "N")
}
```

```{r, eval=FALSE}
set.seed(123)
simil_mat = get_similarity_mat(cluster_num, confuse_rate=0.2)
mis_idx = runif(1, -0.1, 0.1)
change_rate = simil_mat[1,1]*abs(mis_idx)
if (mis_idx < 0){
  simil_mat[1,1] = simil_mat[1,1] - change_rate
  simil_mat[1,3] = simil_mat[1,3] + change_rate
} else {
  simil_mat[3,1] = simil_mat[1,1] - change_rate
  simil_mat[3,3] = simil_mat[1,3] + change_rate
}
```

```{r, eval=FALSE}
## one time example
## Data generation part
set.seed(123)
totals1 = ceiling(runif(rep1, sample_size1, sample_size2))  # number of replicates + sample size in each replicate
totals2 = ceiling(runif(rep2, sample_size1, sample_size2))
diri_s1 = probC1 * concentration  # True proportions + concentration
diri_s2 = probC2 * concentration

## Test part 1
### DCATS
dcats_res1 = betabinLRT_rw(numb_cond1, numb_cond2, bias_corr = FALSE)
### fisher
fisher_pvals1 = getFisher(numb_cond1, numb_cond2)


## Bias Correction
numb_cond1BC = numb_cond1
numb_cond2BC = numb_cond2
for (i in seq_len(nrow(numb_cond1))) {
  numb_cond1BC[i, ] <- sum(numb_cond1[i, ]) *
    multinom_EM(numb_cond1[i, ], simil_mat, verbose = FALSE)$mu
  numb_cond1BC[i, ] = ceiling(numb_cond1BC[i, ])
  }
for (i in seq_len(nrow(numb_cond2))) {
  numb_cond1BC[i, ] <- sum(numb_cond2[i, ]) *
    multinom_EM(numb_cond2[i, ], simil_mat, verbose = FALSE)$mu
  numb_cond1BC[i, ] = ceiling(numb_cond1BC[i, ])
  }

## Test part 2
### DCATS
dcats_res2 = betabinLRT_rw(numb_cond1BC, numb_cond2BC, bias_corr = FALSE)
### fisher
fisher_pvals2 = getFisher(numb_cond1BC, numb_cond2BC)

all_res = data.frame(cluster = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L")[1:cluster_num], truth = truthRes) %>% 
    mutate(dcats_pvals1 = dcats_res1$pvals,
           fisher_pvals1 = fisher_pvals1,
           dcats_pvals2 = dcats_res2$pvals,
           fisher_pvals2 = fisher_pvals2)
```

```{r}
set.seed(123)
simulationDF_list = vector(mode = "list", length = simulation_times)
for (j in 1:simulation_times) {
  simulationDF = data.frame()
  for (i in 1:simulation_size){
    ## create confusion matrix that is not uniformly distributed
    simil_mat = get_similarity_mat(cluster_num, confuse_rate=0.2)
    if (FALSE){
      mis_idx = runif(1, -0.2, 0.2)
      change_rate = simil_mat[1,1]*abs(mis_idx)
      if (mis_idx < 0){
        simil_mat[1,1] = simil_mat[1,1] - change_rate
        simil_mat[1,3] = simil_mat[1,3] + change_rate
        } else {
          simil_mat[3,1] = simil_mat[1,1] - change_rate
          simil_mat[3,3] = simil_mat[1,3] + change_rate
        }
      }
    
    
    ## Data generation part
    totals1 = ceiling(runif(rep1, sample_size1, sample_size2))
    totals2 = ceiling(runif(rep2, sample_size1, sample_size2))
    diri_s1 = probC1 * concentration
    diri_s2 = probC2 * concentration
    sim_dat <- DCATS::simulator_base(totals1, totals2, diri_s1, diri_s2, simil_mat)
    numb_cond1 = as.matrix(sim_dat$numb_cond1)
    numb_cond2 = as.matrix(sim_dat$numb_cond2)
    
    ## Test part 1
    ### DCATS
    dcats_res1 = betabinLRT_rw(numb_cond1, numb_cond2, bias_corr = FALSE)
    ### fisher
    fisher_pvals1 = getFisher(numb_cond1, numb_cond2)
    
    ## Bias Correction
    numb_cond1BC = numb_cond1
    numb_cond2BC = numb_cond2
    for (i in seq_len(nrow(numb_cond1))) {
      numb_cond1BC[i, ] <- sum(numb_cond1[i, ]) *
        multinom_EM(numb_cond1[i, ], simil_mat, verbose = FALSE)$mu
      numb_cond1BC[i, ] = ceiling(numb_cond1BC[i, ])
      }
    for (i in seq_len(nrow(numb_cond2))) {
      numb_cond1BC[i, ] <- sum(numb_cond2[i, ]) *
        multinom_EM(numb_cond2[i, ], simil_mat, verbose = FALSE)$mu
      numb_cond1BC[i, ] = ceiling(numb_cond1BC[i, ])
      }
  
    ## Test part 2
    ### DCATS
    dcats_res2 = betabinLRT_rw(numb_cond1BC, numb_cond2BC, bias_corr = FALSE)
    ### fisher
    fisher_pvals2 = getFisher(numb_cond1BC, numb_cond2BC)
  
    all_res = data.frame(cluster = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L")[1:cluster_num], truth = truthRes) %>% 
      mutate(dcats_noBC = dcats_res1$pvals,
            fisher_noBC = fisher_pvals1,
            dcats_withBC = dcats_res2$pvals,
            fisher_withBC = fisher_pvals2)

    simulationDF = rbind(simulationDF, all_res)
    }
  simulationDF_list[[j]] = simulationDF
}

file_name = str_c("./data/simulationRES/repplicates", as.character(rep1), "&", as.character(rep2), "_K", as.character(cluster_num), "_con", as.character(concentration), "_countSim.RData")
save(simulationDF_list, file = file_name)

## send email after finishing
sendEmail("Simulation is done!!")
```

```{r}
load(file_name)
simulationDF_list[[1]] %>% 
  pivot_longer(dcats_noBC:fisher_withBC,
    names_to = "statistics", 
    values_to = "pvals") %>% 
  ggplot(aes(x = statistics, y = pvals)) +
  geom_boxplot()
```

Calculate PPV NPV TNR F1 Power FOR FNR FPR FDR AUC

```{r, eval=FALSE}
simulationDF = simulationDF_list[[1]]
library(pROC)
methods = colnames(simulationDF)[3:6]
roc_rose <- plot(roc(simulationDF$truth, simulationDF[,3]), legacy.axes = TRUE, print.auc = TRUE, print.auc.y = .7, print.auc.x = .7, col = 1)
graphics::text(0.2, 0.7, paste(methods[1]), col=1)
for (i in 2:4){
  roc_rose <- plot(roc(simulationDF$truth, simulationDF[,i+2]), legacy.axes = TRUE, print.auc = TRUE, print.auc.y = .7-(i-1)*0.07, print.auc.x = .7, col = i, add = TRUE)
  graphics::text(0.2, 0.7-(i-1)*0.07, paste(methods[i]), col=i)
}

```

```{r, eval=FALSE}
simulationDF = simulationDF %>% 
  mutate(dcats_tf1 = ifelse(dcats_pvals1 < 0.05, "P", "N"),
         fisher_tf1 = ifelse(fisher_pvals1 < 0.05, "P", "N"),
         dcats_tf2 = ifelse(dcats_pvals2 < 0.05, "P", "N"),
         fisher_tf2 = ifelse(fisher_pvals2 < 0.05, "P", "N"))

evaluationDF = data.frame()
methods = colnames(simulationDF)[3:6]
numb_mthd = length(methods)
auc = rep(NA, numb_mthd)
recall = rep(NA, numb_mthd)
FNR = rep(NA, numb_mthd)
FPR = rep(NA, numb_mthd)
specificity = rep(NA, numb_mthd)
precision = rep(NA, numb_mthd)
FDR = rep(NA, numb_mthd)
FOR = rep(NA, numb_mthd)
NPV = rep(NA, numb_mthd)
accuracy = rep(NA, numb_mthd)
F1 = rep(NA, numb_mthd)

for (i in 3:6){
  auc[i-2] = roc(simulationDF$truth, simulationDF[,i])$auc
  truth = simulationDF$truth
  pred = simulationDF[, i+4]
  confu_mtr = table(pred, truth)
  TN = confu_mtr[1,1] 
  FN = confu_mtr[1,2]
  FP = confu_mtr[2,1]
  TP = confu_mtr[2,2]
  truthN = TN + FP
  truthP = FN + TP
  predP = TP + FP
  predN = FN + TN
  recall[i-2] = TP/truthP
  FNR[i-2] = FN/truthP
  FPR[i-2] = FP/truthN
  specificity[i-2] = TN/truthN
  precision[i-2] = TP/predP
  FDR[i-2] = FP/predP
  FOR[i-2] = FN/predN
  NPV[i-2] = TN/predN
  accuracy[i-2] = (TP+TN)/(truthP+truthN)
  F1[i-2] = 2*TP/(2*TP+FP+FN)
}

res = data.frame(methods = methods, auc = auc, recall = recall, FNR = FNR, FPR = FPR, specificity = specificity, precision = precision, FDR = FDR, FOR = FOR, NPV = NPV, accuracy = accuracy, F1 = F1)
```

```{r}
evaluationDF = data.frame()
len = length(simulationDF_list)
for (j in 1:len){
  simulationDF = simulationDF_list[[j]] %>% 
  mutate(dcats_tf1 = ifelse(dcats_noBC < 0.05, "P", "N"),
         fisher_tf1 = ifelse(fisher_noBC < 0.05, "P", "N"),
         dcats_tf2 = ifelse(dcats_withBC < 0.05, "P", "N"),
         fisher_tf2 = ifelse(fisher_withBC < 0.05, "P", "N"))
  methods = colnames(simulationDF)[3:6]
  numb_mthd = length(methods)
  auc = rep(NA, numb_mthd)
  recall = rep(NA, numb_mthd)
  FNR = rep(NA, numb_mthd)
  FPR = rep(NA, numb_mthd)
  specificity = rep(NA, numb_mthd)
  precision = rep(NA, numb_mthd)
  FDR = rep(NA, numb_mthd)
  FOR = rep(NA, numb_mthd)
  NPV = rep(NA, numb_mthd)
  accuracy = rep(NA, numb_mthd)
  F1 = rep(NA, numb_mthd)
  for (i in 3:6){
    auc[i-2] = roc(simulationDF$truth, simulationDF[,i])$auc
    truth = simulationDF$truth
    pred = simulationDF[, i+4]
    confu_mtr = table(pred, truth)
    if(dim(confu_mtr)[1] + dim(confu_mtr)[2] == 4){
      TN = confu_mtr[1,1] 
      FN = confu_mtr[1,2]
      FP = confu_mtr[2,1]
      TP = confu_mtr[2,2]
      truthN = TN + FP
      truthP = FN + TP
      predP = TP + FP
      predN = FN + TN
      recall[i-2] = TP/truthP
      FNR[i-2] = FN/truthP
      FPR[i-2] = FP/truthN
      specificity[i-2] = TN/truthN
      precision[i-2] = TP/predP
      FDR[i-2] = FP/predP
      FOR[i-2] = FN/predN
      NPV[i-2] = TN/predN
      accuracy[i-2] = (TP+TN)/(truthP+truthN)
      F1[i-2] = 2*TP/(2*TP+FP+FN)
      }
  res = data.frame(trial = as.character(j), methods = methods, auc = auc, recall = recall, FNR = FNR, FPR = FPR, specificity = specificity, precision = precision, FDR = FDR, FOR = FOR, NPV = NPV, accuracy = accuracy, F1 = F1)
  evaluationDF = rbind(evaluationDF, res)
    }
    
}

evaluationDF %>% 
  tidyr::separate(methods, c("method", "condition"), sep = "_") %>% 
  pivot_longer(
    auc:F1,
    names_to = "statistics", 
    values_to = "value") %>% 
  ggplot(aes(x = statistics, y = value, color = method)) + geom_boxplot() + facet_grid(.~condition) +
  theme(axis.text.x = element_text(angle = 45))
```

## Previous trail

Add more misclassification manually

```{r}
set.seed(123)
K = 3
totals1 = ceiling(runif(4, 500, 2000))
totals2 = ceiling(runif(2, 500, 2000))
diri_s1 = c(1/3, 1/3, 1/3) * 20
diri_s2 = c(1/3, 1/2, 1/6) * 20
simil_mat = get_similarity_mat(K, confuse_rate=0.1)
sim_dat <- DCATS::simulator_base(totals1, totals2, diri_s1, diri_s2, simil_mat)
print(sim_dat)
```

Add misclassification between cluster 1 and cluster 3

```{r}
rate1 = runif(dim(sim_dat$numb_cond1)[1], -0.2, 0.2)
change_num = (sim_dat$numb_cond1[,1]*rate1) %>% round()
sim_dat$numb_cond1[,1] = sim_dat$numb_cond1[,1] - change_num
sim_dat$numb_cond1[,3] = sim_dat$numb_cond1[,3] + change_num
print(sim_dat$numb_cond1)

rate2 = runif(dim(sim_dat$numb_cond2)[1], -0.2, 0.2)
change_num = (sim_dat$numb_cond2[,1]*rate2) %>% round()
sim_dat$numb_cond2[,1] = sim_dat$numb_cond2[,1] - change_num
sim_dat$numb_cond2[,3] = sim_dat$numb_cond2[,3] + change_num
print(sim_dat$numb_cond2)

change_rate = (sum(rate1*sim_dat$numb_cond1[,1]) + sum(rate2*sim_dat$numb_cond2[,2]))/(sum(sim_dat$numb_cond1[,1]) + sum(sim_dat$numb_cond2[,1]))
if (change_rate > 0){
  simil_matT = rbind(c(1-change_rate, 0, change_rate), c(0, 1, 0), c(0, 0, 1))
} else {
  simil_matT = rbind(c(1, 0, 0), c(0, 1, 0), c(-change_rate, 0, 1+change_rate))
}
print(simil_matT)
```

```{r}
## DCATS---betabinLRT
  betabin_noBC = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = FALSE)
  ##  DCATS---betabinLRT with Bias correction from similarity matrix
  simil_matI = get_similarity_mat(K, confuse_rate=0)
  simil_matU = get_similarity_mat(K, confuse_rate=0.1)
  betabin_wBC_I = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = TRUE, similarity_mat = simil_matI)
  betabin_wBC_U = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = TRUE, similarity_mat = simil_matU)
  betabin_wBC_T = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = TRUE, similarity_mat = simil_matT)
  ## DCATS---betabin with similarity matrix (backward-forward)
  betabin_wMI = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matI, n_samples = 100, binom_only = FALSE)
  betabin_wMU = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matU, n_samples = 100, binom_only = FALSE)
  betabin_wMT = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matT, n_samples = 100, binom_only = FALSE)
  bin_wMI = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matI, n_samples = 100)
  bin_wMU = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matU, n_samples = 100)
  bin_wMT = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matT, n_samples = 100)
  ## Fisher's exact test
  fisher_pvals = getFisher(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2))
  all_res = data.frame(cluster = c("A", "B", "C"), truth = c("N", "P", "P")) %>% 
    mutate(betabin_noBC_pvals = betabin_noBC$pvals,
           betabin_wBCI_pvals = betabin_wBC_I$pvals,
           betabin_wBCU_pvals = betabin_wBC_U$pvals,
           betabin_wBCT_pvals = betabin_wBC_T$pvals,
           betabin_wMI_pvals = betabin_wMI$pvals,
           betabin_wMU_pvals = betabin_wMU$pvals,
           betabin_wMT_pvals = betabin_wMT$pvals,
           bin_wMI_pvals = bin_wMI$pvals,
           bin_wMU_pvals = bin_wMU$pvals,
           bin_wMT_pvals = bin_wMT$pvals,
           fisher_pvals = fisher_pvals)
  print(all_res)
```

### 3 clusters

```{r}
set.seed(123)
K = 3
totals1 = ceiling(runif(4, 500, 2000))
totals2 = ceiling(runif(2, 500, 2000))
diri_s1 = c(1/3, 1/3, 1/3) * 20
diri_s2 = c(1/3, 1/2, 1/6) * 20
simil_mat = get_similarity_mat(K, confuse_rate=0.2)
sim_dat <- DCATS::simulator_base(totals1, totals2, diri_s1, diri_s2, simil_mat)
```

```{r}
simulationDF = data.frame()
for (i in 1:20){
  simil_mat = get_similarity_mat(K, confuse_rate=0.2)
  sim_dat <- DCATS::simulator_base(totals1, totals2, diri_s1, diri_s2, simil_mat)
  
  ## DCATS---betabinLRT
  betabin_noBC = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = FALSE)
  ##  DCATS---betabinLRT with Bias correction from similarity matrix
  simil_matI = get_similarity_mat(K, confuse_rate=0)
  betabin_wBC_I = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = TRUE, similarity_mat = simil_matI)
  betabin_wBC_T = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = TRUE, similarity_mat = simil_mat)
  ## DCATS---betabin with similarity matrix (backward-forward)
  betabin_wMI = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matI, n_samples = 100, binom_only = FALSE)
  betabin_wMT = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_mat, n_samples = 100, binom_only = FALSE)
  bin_wMI = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matI, n_samples = 100)
  bin_wMT = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_mat, n_samples = 100)
  ## Fisher's exact test
  fisher_pvals = getFisher(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2))
  all_res = data.frame(cluster = c("A", "B", "C"), truth = c("N", "P", "P")) %>% 
    mutate(betabin_noBC_pvals = betabin_noBC$pvals,
           betabin_wBCI_pvals = betabin_wBC_I$pvals,
           betabin_wBCT_pvals = betabin_wBC_T$pvals,
           betabin_wMI_pvals = betabin_wMI$pvals,
           betabin_wMT_pvals = betabin_wMT$pvals,
           bin_wMI_pvals = bin_wMI$pvals,
           bin_wMT_pvals = bin_wMT$pvals,
           fisher_pvals = fisher_pvals)
  simulationDF = rbind(simulationDF, all_res)
}
head(simulationDF)

## send email after finishing
sendEmail("Simulation is done!!")
```

```{r}
library(pROC)
methods = colnames(simulationDF)[3:10]
roc_rose <- plot(roc(simulationDF$truth, simulationDF[,3]), legacy.axes = TRUE, print.auc = TRUE, print.auc.y = .7, print.auc.x = .7, col = 1)
graphics::text(0.2, 0.7, paste(methods[1]), col=1)
for (i in 2:8){
  roc_rose <- plot(roc(simulationDF$truth, simulationDF[,i+2]), legacy.axes = TRUE, print.auc = TRUE, print.auc.y = .7-(i-1)*0.07, print.auc.x = .7, col = i, add = TRUE)
  graphics::text(0.2, 0.7-(i-1)*0.07, paste(methods[i]), col=i)
}
```

### 7 clusters

```{r}
set.seed(123)
K = 7
totals1 = ceiling(runif(4, 1000, 2000))
totals2 = ceiling(runif(2, 1000, 2000))
diri_s1 = c(rep(0.1, 4), rep(0.2, 3)) * 50
diri_s2 = c(0.1, 0.1, 0.2, 0.05, 0.2, 0.3, 0.05) * 50
```

```{r}
set.seed(123)
simulationDF = data.frame()
for (i in 1:50){
  simil_mat = get_similarity_mat(K, confuse_rate=0.4)
  sim_dat <- DCATS::simulator_base(totals1, totals2, diri_s1, diri_s2, simil_mat)
  
  ## DCATS---betabinLRT
  betabin_noBC = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = FALSE)
  ##  DCATS---betabinLRT with Bias correction from similarity matrix
  simil_matI = get_similarity_mat(K, confuse_rate=0)
  betabin_wBC_I = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = TRUE, similarity_mat = simil_matI)
  betabin_wBC_T = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = TRUE, similarity_mat = simil_mat)
  ## DCATS---betabin with similarity matrix (backward-forward)
  betabin_wMI = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matI, n_samples = 100, binom_only = FALSE)
  betabin_wMT = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_mat, n_samples = 100, binom_only = FALSE)
  bin_wMI = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matI, n_samples = 100)
  bin_wMT = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_mat, n_samples = 100)
  ## Fisher's exact test
  fisher_pvals = getFisher(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2))
  all_res = data.frame(cluster = c("A", "B", "C", "D", "E", "F", "G"), truth = c("N", "N", "P", "P", "N", "P", "P")) %>% 
    mutate(betabin_noBC_pvals = betabin_noBC$pvals,
           betabin_wBCI_pvals = betabin_wBC_I$pvals,
           betabin_wBCT_pvals = betabin_wBC_T$pvals,
           betabin_wMI_pvals = betabin_wMI$pvals,
           betabin_wMT_pvals = betabin_wMT$pvals,
           bin_wMI_pvals = bin_wMI$pvals,
           bin_wMT_pvals = bin_wMT$pvals,
           fisher_pvals = fisher_pvals)
  simulationDF = rbind(simulationDF, all_res)
}
head(simulationDF)

## send email after finishing
sendEmail("Simulation is done!!")
```

```{r}
library(pROC)
methods = colnames(simulationDF)[3:10]
roc_rose <- plot(roc(simulationDF$truth, simulationDF[,3]), legacy.axes = TRUE, print.auc = TRUE, print.auc.y = .7, print.auc.x = .5, col = 1)
graphics::text(0.1, 0.68, paste(methods[1]), col=1)
for (i in 2:8){
  roc_rose <- plot(roc(simulationDF$truth, simulationDF[,i+2]), legacy.axes = TRUE, print.auc = TRUE, print.auc.y = .7-(i-1)*0.07, print.auc.x = .5, col = i, add = TRUE)
  graphics::text(0.1, 0.68-(i-1)*0.07, paste(methods[i]), col=i)
}
```

### 3 clusters(not uniform misclassfy)

```{r}
set.seed(123)
K = 3  # number of cluster
totals1 = ceiling(runif(7, 500, 2000))  # number of replicates + sample size in each replicate
totals2 = ceiling(runif(6, 500, 2000))
diri_s1 = c(1/3, 1/3, 1/3) * 20  # True proportions + concentration
diri_s2 = c(1/3, 1/2, 1/6) * 20
simil_mat = get_similarity_mat(K, confuse_rate=0.2)  # similarity matrix used in simulator to introduce misclassification
```

```{r}
simulationDF = data.frame()
for (i in 1:50){
  sim_dat <- DCATS::simulator_base(totals1, totals2, diri_s1, diri_s2, simil_mat)
  ## introduce more misclassification manually
  rate1 = runif(dim(sim_dat$numb_cond1)[1], -0.2, 0.2)
  change_num = (sim_dat$numb_cond1[,1]*rate1) %>% round()
  sim_dat$numb_cond1[,1] = sim_dat$numb_cond1[,1] - change_num
  sim_dat$numb_cond1[,3] = sim_dat$numb_cond1[,3] + change_num
  
  rate2 = runif(dim(sim_dat$numb_cond2)[1], -0.2, 0.2)
  change_num = (sim_dat$numb_cond2[,1]*rate2) %>% round()
  sim_dat$numb_cond2[,1] = sim_dat$numb_cond2[,1] - change_num
  sim_dat$numb_cond2[,3] = sim_dat$numb_cond2[,3] + change_num
  ## calculate true matrix
  change_rate = (sum(rate1*sim_dat$numb_cond1[,1]) + sum(rate2*sim_dat$numb_cond2[,2]))/(sum(sim_dat$numb_cond1[,1]) + sum(sim_dat$numb_cond2[,1]))
  if (change_rate > 0){
    simil_matT = rbind(c(1-change_rate, 0, change_rate), c(0, 1, 0), c(0, 0, 1))
} else {
  simil_matT = rbind(c(1, 0, 0), c(0, 1, 0), c(-change_rate, 0, 1+change_rate))
}
  ## DCATS---betabinLRT
  betabin_noBC = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = FALSE)
  ##  DCATS---betabinLRT with Bias correction from similarity matrix
  simil_matI = get_similarity_mat(K, confuse_rate=0)
  simil_matU = get_similarity_mat(K, confuse_rate=0.1)
  betabin_wBC_I = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = TRUE, similarity_mat = simil_matI)
  betabin_wBC_U = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = TRUE, similarity_mat = simil_matU)
  betabin_wBC_T = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = TRUE, similarity_mat = simil_matT)
  ## DCATS---betabin with similarity matrix (backward-forward)
  betabin_wMI = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matI, n_samples = 100, binom_only = FALSE)
  betabin_wMU = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matU, n_samples = 100, binom_only = FALSE)
  betabin_wMT = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matT, n_samples = 100, binom_only = FALSE)
  bin_wMI = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matI, n_samples = 100)
  bin_wMU = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matU, n_samples = 100)
  bin_wMT = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matT, n_samples = 100)
  ## Fisher's exact test
  fisher_pvals = getFisher(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2))
  all_res = data.frame(cluster = c("A", "B", "C"), truth = c("N", "P", "P")) %>% 
    mutate(betabin_noBC_pvals = betabin_noBC$pvals,
           betabin_wBCI_pvals = betabin_wBC_I$pvals,
           betabin_wBCU_pvals = betabin_wBC_U$pvals,
           betabin_wBCT_pvals = betabin_wBC_T$pvals,
           betabin_wMI_pvals = betabin_wMI$pvals,
           betabin_wMU_pvals = betabin_wMU$pvals,
           betabin_wMT_pvals = betabin_wMT$pvals,
           bin_wMI_pvals = bin_wMI$pvals,
           bin_wMU_pvals = bin_wMU$pvals,
           bin_wMT_pvals = bin_wMT$pvals,
           fisher_pvals = fisher_pvals)
  simulationDF = rbind(simulationDF, all_res)
}
head(simulationDF)
```

```{r}
library(pROC)
methods = colnames(simulationDF)[3:13]
roc_rose <- plot(roc(simulationDF$truth, simulationDF[,3]), legacy.axes = TRUE, print.auc = TRUE, print.auc.y = .7, print.auc.x = .7, col = 1)
graphics::text(0.2, 0.7, paste(methods[1]), col=1)
for (i in 2:11){
  roc_rose <- plot(roc(simulationDF$truth, simulationDF[,i+2]), legacy.axes = TRUE, print.auc = TRUE, print.auc.y = .7-(i-1)*0.07, print.auc.x = .7, col = i, add = TRUE)
  graphics::text(0.2, 0.7-(i-1)*0.07, paste(methods[i]), col=i)
}
```

### 7 clusters

```{r}
set.seed(123)
K = 7
totals1 = ceiling(runif(4, 1000, 2500))
totals2 = ceiling(runif(2, 1000, 2500))
diri_s1 = c(rep(0.1, 4), rep(0.2, 3)) * 20
diri_s2 = c(0.1, 0.1, 0.2, 0.05, 0.2, 0.3, 0.05) * 20
```

```{r}
set.seed(123)
simulationDF = data.frame()
for (i in 1:20){
  simil_mat = get_similarity_mat(K, confuse_rate=0.2)
  change_rate = runif(1, -0.2, 0.2)
  if (change_rate < 0){
    simil_mat[5, 5] = simil_mat[5, 5](1+change_rate)
    simil_mat[]
  }
  sim_dat <- DCATS::simulator_base(totals1, totals2, diri_s1, diri_s2, simil_mat)
  ## get count table
  rate1 = runif(dim(sim_dat$numb_cond1)[5], -0.2, 0.2)
  change_num = (sim_dat$numb_cond1[,5]*rate1) %>% round()
  sim_dat$numb_cond1[,5] = sim_dat$numb_cond1[,5] - change_num
  sim_dat$numb_cond1[,6] = sim_dat$numb_cond1[,6] + change_num
  
  rate2 = runif(dim(sim_dat$numb_cond2)[5], -0.2, 0.2)
  change_num = (sim_dat$numb_cond2[,5]*rate2) %>% round()
  sim_dat$numb_cond2[,5] = sim_dat$numb_cond2[,5] - change_num
  sim_dat$numb_cond2[,6] = sim_dat$numb_cond2[,6] + change_num
  ## calculate true matrix
  change_rate = (sum(rate1*sim_dat$numb_cond1[,5]) + sum(rate2*sim_dat$numb_cond2[,6]))/(sum(sim_dat$numb_cond1[,5]) + sum(sim_dat$numb_cond2[,6]))
  if (change_rate > 0){
    simil_matT = rbind(c(1,0,0,0,0,0,0,0)c(1-change_rate, 0, change_rate), c(0, 1, 0), c(0, 0, 1))
} else {
  simil_matUT = rbind(c(1, 0, 0), c(0, 1, 0), c(-change_rate, 0, 1+change_rate))
}
  ## DCATS---betabinLRT
  betabin_noBC = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = FALSE)
  ##  DCATS---betabinLRT with Bias correction from similarity matrix
  simil_matI = get_similarity_mat(K, confuse_rate=0)
  simil_matU = get_similarity_mat(K, confuse_rate=0.1)
  betabin_wBC_I = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = TRUE, similarity_mat = simil_matI)
  betabin_wBC_U = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = TRUE, similarity_mat = simil_matU)
  betabin_wBC_T = betabinLRT_rw(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), bias_corr = TRUE, similarity_mat = simil_matT)
  ## DCATS---betabin with similarity matrix (backward-forward)
  betabin_wMI = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matI, n_samples = 100, binom_only = FALSE)
  betabin_wMU = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matU, n_samples = 100, binom_only = FALSE)
  betabin_wMT = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matT, n_samples = 100, binom_only = FALSE)
  bin_wMI = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matI, n_samples = 100)
  bin_wMU = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matU, n_samples = 100)
  bin_wMT = dcats_betabin(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2), similarity_mat = simil_matT, n_samples = 100)
  ## Fisher's exact test
  fisher_pvals = getFisher(as.matrix(sim_dat$numb_cond1), as.matrix(sim_dat$numb_cond2))
  all_res = data.frame(cluster = c("A", "B", "C"), truth = c("N", "N", "P", "P", "N", "P", "P")) %>% 
    mutate(betabin_noBC_pvals = betabin_noBC$pvals,
           betabin_wBCI_pvals = betabin_wBC_I$pvals,
           betabin_wBCU_pvals = betabin_wBC_U$pvals,
           betabin_wBCT_pvals = betabin_wBC_T$pvals,
           betabin_wMI_pvals = betabin_wMI$pvals,
           betabin_wMU_pvals = betabin_wMU$pvals,
           betabin_wMT_pvals = betabin_wMT$pvals,
           bin_wMI_pvals = bin_wMI$pvals,
           bin_wMU_pvals = bin_wMU$pvals,
           bin_wMT_pvals = bin_wMT$pvals,
           fisher_pvals = fisher_pvals)
  simulationDF = rbind(simulationDF, all_res)
}
head(simulationDF)
```

```{r, message=FALSE}
library(pROC)
methods = colnames(simulationDF)[3:13]
roc_rose <- plot(roc(simulationDF$truth, simulationDF[,3]), legacy.axes = TRUE, print.auc = TRUE, print.auc.y = .7, print.auc.x = .5, col = 1)
graphics::text(0.05, 0.69, paste(methods[1]), col=1)
for (i in 2:11){
  roc_rose <- plot(roc(simulationDF$truth, simulationDF[,i+2]), legacy.axes = TRUE, print.auc = TRUE, print.auc.y = .7-(i-1)*0.07, print.auc.x = .5, col = i, add = TRUE)
  graphics::text(0.05, 0.69-(i-1)*0.07, paste(methods[i]), col=i)
}

print("Time")
apply(na.omit(time), 2, mean)
```