---
title: "Simulation"
author: "Xinyi Lin"
date: "9/10/2020"
output:
  pdf_document: default
  html_document: default
---

Problems need to solve:

3. DCATS: n_samples? do p-values correspond to each cluster? 

```{r,message=FALSE,warning=FALSE}
library(splatter)
library(Seurat)
library(speckle)
library(DCATS)
library(ggplot2)
library(tidyverse)
library(scdney)
library(diffcyt)
```

## Simulation 1

Set up the group probabilities for Normal and Mutate conditions. Need to change resolution as well.

```{r}
probNor = c(1/3,1/3,1/3)
probMut = c(1/2,1/6,1/3)
setresolu = 0.5
```

### Simulation

First, we simulate data with three groups and two of them are similar, which means they might have high misclustering rate. 'de.prob' specifies the probability that a gene selected is differentially expressed between the cluster and the rest of the cells.

```{r}
set.seed(12345)

# simulate normal
param.groups <- newSplatParams(batchCells = c(600, 600, 600), nGenes = 100)
simNor <- splatSimulateGroups(param.groups, group.prob = probNor, de.prob = c(0.1,0.1,0.5), verbose = FALSE)
simNor@colData@rownames = str_replace(simNor@colData@rownames, "Cell", "NorCell")
simNor_mat <- counts(simNor)

# simulate mutate
simMut <- splatSimulateGroups(param.groups, group.prob = probMut, de.prob = c(0.1,0.1,0.5), verbose = FALSE)
simMut@colData@rownames = str_replace(simMut@colData@rownames, "Cell", "MutCell")
simMut_mat <- counts(simMut)
```

Batch information of normal and mutate sample

```{r}
batchNor = simNor@colData@listData$Batch %>% 
  str_replace("Batch", "Nor")
batchMut = simMut@colData@listData$Batch %>% 
  str_replace("Batch", "Mut")
```

Then, separate the normal sample and mutate sample by batch and combine them into a list.

```{r,warning=FALSE,message=FALSE}
# Normal
seuratNor <- CreateSeuratObject(counts = simNor_mat, project="Splatter")
seuratNor <- AddMetaData(object = seuratNor, metadata = batchNor, col.name = 'batch')
seuratNor <- AddMetaData(object = seuratNor, metadata = rep("Normal", 1800), col.name = 'condition')
# Mutate
seuratMut <- CreateSeuratObject(counts = simMut_mat, project="Splatter")
seuratMut <- AddMetaData(object = seuratMut, metadata = batchMut, col.name = 'batch')
seuratMut <- AddMetaData(object = seuratMut, metadata = rep("Mutate", 1800), col.name = 'condition')
```

```{r}
# split by batch
listNor = SplitObject(seuratNor, split.by = "batch")
listMut = SplitObject(seuratMut, split.by = "batch")

# combine Normal and Mutate
listSamples = c(listNor, listMut)
```

### Process by using Seurat

Log-normalization and identify variable features for each batches separately

```{r}
for (i in 1:length(listSamples)) {
    listSamples[[i]] <- NormalizeData(listSamples[[i]], verbose = FALSE)
    listSamples[[i]] <- FindVariableFeatures(listSamples[[i]], selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
}
```

Integrate all batches

```{r}
anchors <- FindIntegrationAnchors(object.list = listSamples, dims = 1:30, verbose = FALSE)
integratedSamples <- IntegrateData(anchorset = anchors, dims = 1:30, verbose = FALSE)
```

```{r}
DefaultAssay(integratedSamples) <- "integrated"

# Run the standard workflow for visualization and clustering
integratedSamples <- ScaleData(integratedSamples, verbose = FALSE)
integratedSamples <- RunPCA(integratedSamples, npcs = 30, verbose = FALSE)
```

```{r, eval = FALSE}
# check how many PCs to choose
ElbowPlot(integratedSamples)
# dim = 5
```

```{r}
integratedSamples<-FindNeighbors(integratedSamples, dims = 1:5, verbose = FALSE)
integratedSamples<-FindClusters(integratedSamples, resolution = setresolu, algorithm=2, verbose = FALSE)
```

```{r}
integratedSamples <- RunUMAP(integratedSamples, reduction = "pca", dims = 1:30, verbose = FALSE)
```

```{r}
integratedSamples@active.ident = integratedSamples@active.ident %>% 
  plyr::mapvalues(from = c("1", "0", "2"), to = c("A", "B", "C"))
# the plot which shows the cluster results of different samples
DimPlot(integratedSamples, reduction = "umap", group.by = "batch") + ggtitle("Cluster Results of Different Samples")
# the plot which shows the cluster results of different clusters
DimPlot(integratedSamples, reduction = "umap") + ggtitle("Cluster Results")
# the plot which shows the cluster results of different clusters in different samples
DimPlot(integratedSamples, ncol = 3, reduction = "umap", split.by = "batch") + ggtitle("Cluster Results of Different Clusters in Different Samples")
```

### Confusion matrix

```{r}
celllabels_orig = c(simNor@colData@listData$Group, simMut@colData@listData$Group) 

conf.mat<-table(Idents(integratedSamples), celllabels_orig)
print(conf.mat)

true.conf<-t(t(conf.mat)/apply(conf.mat,2,sum))

print(true.conf)
```

```{r}
condition = integratedSamples@meta.data$condition
condNor<-Idents(integratedSamples)[condition == "Normal"];
condMut<-Idents(integratedSamples)[condition == "Mutate"];
```

### Fisher's exact test

```{r}
dfRes = data.frame(clusterRes = integratedSamples@active.ident, batch = integratedSamples$batch, condition = integratedSamples$condition) %>% 
  tibble::rownames_to_column("cellID")
head(dfRes)
```

```{r}
dfCount = dfRes %>% 
  group_by(condition, clusterRes) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = "clusterRes", values_from = "n") %>% 
  mutate(nonA = B + C,
         nonB = A + C,
         nonC = A + B) %>% 
  select(A, B, C:nonC)

print(dfCount)
```

```{r, eval=FALSE, echo=FALSE}
fisherP = rep(NA,3)
for (i in 1:3){
  fisherP[i] = fisher.test(dfCount[,c(i+1,i+4)])$p.value
}

fisherRes = data.frame(clust = c("A", "B", "C"), p_value = fisherP)
print(fisherRes)
```

### speckle

```{r}
# clusters indicates the cluster results, sample indicates the biological replicates, group indicates the conditions/groups
propeller(clusters = dfRes$clusterRes, sample = dfRes$batch, group = dfRes$condition)

# Plot cell type proportions
#plotCellTypeProps(clusters=speckleData$clusterRes, sample=speckleData$batch)
```

### DCATS

First, get the count tables for normal and mutate condition. The count tables count numbers of cells in different clusters. This is `counts1` and `counts2`. The `similarity_mat` is calculated for all batches across different condtions. It is the true.conf we get before.

*how to set the n_samples parameter?*

```{r}
countNor = table(batchNor, condNor)
countMut = table(batchMut, condMut)
dcats_fit(countNor, countMut, true.conf)
```

### scDC

scDC doesn't consider the difference between different conditions and different batches.

```{r}
res_scDC <- scDC_noClustering(cellTypes = dfRes$clusterRes, dfRes$batch, calCI = TRUE, 
                                     calCI_method = c("percentile", "BCa", "multinom"),
                                     nboot = 50)
```

```{r}
barplotCI(res_scDC, c(rep("cond1",9),rep("cond2",9)))
```

```{r,eval=FALSE}
source("glm.R")
res_GLM <- fitGLM(res_scDC, c(rep("cond1",9),rep("cond2",9)), pairwise = FALSE)
```

### diffcyt

Create a template object to be replaced in the future

```{r}
# For a complete workflow example demonstrating each step in the 'diffcyt' pipeline, 
# see the package vignette.

# Function to create random data (one sample)
d_random <- function(n = 20000, mean = 0, sd = 1, ncol = 20, cofactor = 5) {
  d <- sinh(matrix(rnorm(n, mean, sd), ncol = ncol)) * cofactor
  colnames(d) <- paste0("marker", sprintf("%02d", 1:ncol))
  d
}

# Create random data (without differential signal)
set.seed(123)
d_input <- list(
  Nor1 = d_random(), 
  Nor2 = d_random(), 
  Nor3 = d_random(), 
  Mut1 = d_random(),
  Mut2 = d_random(),
  Mut3 = d_random()
)

experiment_info <- data.frame(
  sample_id = factor(c("Nor1", "Nor2", "Nor3", "Mut1", "Mut2", "Mut3")), 
  group_id = factor(c(rep("Normal", 3), rep("Mutate", 3))), 
  stringsAsFactors = FALSE
)

marker_info <- data.frame(
  channel_name = paste0("channel", sprintf("%03d", 1:20)), 
  marker_name = paste0("marker", sprintf("%02d", 1:20)), 
  marker_class = factor(c(rep("type", 10), rep("state", 10)), 
                        levels = c("type", "state", "none")), 
  stringsAsFactors = FALSE
)

# Prepare data
d_se <- prepareData(d_input, experiment_info, marker_info)
```

```{r}
# Transform data
d_se <- transformData(d_se)
# Generate clusters
# note: include random seed for reproducible clustering
d_se <- generateClusters(d_se, seed_clustering = 123, xdim = 1, ydim = 3)
# Calculate cluster cell counts
d_counts <- calcCounts(d_se)
```
```{r}

```

