---
title: "tools example"
author: "Xinyi Lin"
date: "9/10/2020"
output: html_document
---

## speckle

```{r,message=FALSE,warning=FALSE}
library(speckle)
library(limma)
library(ggplot2)
library(tidyverse)
```

>In order to test for differences in cell type proportions between multiple experimental conditions at least one of the groups must have some form of biological replication (i.e. at least two samples). For a two group scenario, the absolute minimum sample size is thus three. Since there are many technical aspects which can affect cell type proportion estimates, having biological replication is essential for a meaningful analysis.

```{r,message=FALSE}
# Get some example data which has two groups, three cell types and two 
# biological replicates in each group
cell_info <- speckle_example_data()
head(cell_info)
cell_info %>% 
  group_by(clusters, samples, group) %>% 
  summarise(n = n())
```

```{r}
# Run propeller testing for cell type proportion differences between the two 
# groups
propeller(clusters = cell_info$clusters, sample = cell_info$samples, 
group = cell_info$group)

# Plot cell type proportions
plotCellTypeProps(clusters=cell_info$clusters, sample=cell_info$samples)
```

create my own examples

```{r}
myExample = data.frame(clusters = cell_info$clusters, groups = cell_info$group)
head(myExample)
propeller(clusters = myExample$clusters, group = myExample$groups)
```

## diffcyt

>The diffcyt package implements statistical methods for differential discovery analyses in high-dimensional cytometry data, based on a combination of high-resolution clustering and empirical Bayes moderated tests adapted from transcriptomics.

```{r,warning=FALSE}
suppressPackageStartupMessages(library(HDCytoData))
suppressPackageStartupMessages(library(flowCore))
```

Load Data

```{r}
# d_flowSet
d_flowSet <- Bodenmiller_BCR_XL_flowSet()
print(d_flowSet)
```

what this `d_flowSet` looks like

```{r}
# sample names
pData(d_flowSet)
# number of cells
fsApply(d_flowSet, nrow)
# dimensions
dim(exprs(d_flowSet[[1]]))
# expression values
exprs(d_flowSet[[1]])[1:6, 1:5]
```

```{r}
# Meta-data: experiment information

# check sample order
filenames <- as.character(pData(d_flowSet)$name)

# sample information
sample_id <- gsub("^PBMC8_30min_", "", gsub("\\.fcs$", "", filenames))
group_id <- factor(
  gsub("^patient[0-9]+_", "", sample_id), levels = c("Reference", "BCR-XL")
)
patient_id <- factor(gsub("_.*$", "", sample_id))

experiment_info <- data.frame(
  group_id, patient_id, sample_id, stringsAsFactors = FALSE
)

print(experiment_info)
```

```{r}
# Meta-data: marker information

# source: Bruggner et al. (2014), Table 1

# column indices of all markers, lineage markers, and functional markers
cols_markers <- c(3:4, 7:9, 11:19, 21:22, 24:26, 28:31, 33)
cols_lineage <- c(3:4, 9, 11, 12, 14, 21, 29, 31, 33)
cols_func <- setdiff(cols_markers, cols_lineage)

# channel and marker names
channel_name <- colnames(d_flowSet)
marker_name <- gsub("\\(.*$", "", channel_name)

# marker classes
# note: using lineage markers for 'cell type', and functional markers for 
# 'cell state'
marker_class <- rep("none", ncol(d_flowSet[[1]]))
marker_class[cols_lineage] <- "type"
marker_class[cols_func] <- "state"
marker_class <- factor(marker_class, levels = c("type", "state", "none"))

marker_info <- data.frame(
  channel_name, marker_name, marker_class, stringsAsFactors = FALSE
)

print(marker_info)
```

## DCATS

```{r}
K <- 2
totals1 = c(100, 800, 1300, 600)
totals2 = c(250, 700, 1100)
diri_s1 = rep(1, K) * 20
diri_s2 = rep(1, K) * 20
confuse_rate = 0.2
simil_mat = diag(K) * (1 - confuse_rate) + confuse_rate * (1 - diag(K)) / 3
sim_dat <- DCATS::simulator_base(totals1, totals2, diri_s1, diri_s2, simil_mat)
dcats_fit(sim_dat[[1]], sim_dat[[2]], confuse_mat, n_samples = 100)
```

## scDC

```{r}
library(scdney)
```

```{r}
data("sim") # sample data in the vignette
exprsMat <- sim$sim_exprsMat
subject <- sim$sim_subject
cellTypes <- sim$sim_cellTypes
cond <- sim$sim_cond
```

```{r}
dim(exprsMat)
# 500 genes 260 cells
table(subject, cellTypes)
table(cond, cellTypes)
```

Perform scDC (without clustering)

```{r}
res_scDC_noClust <- scDC_noClustering(cellTypes, subject, calCI = TRUE, 
                                     calCI_method = c("percentile", "BCa", "multinom"),
                                     nboot = 50)
```
Visualization

```{r}
barplotCI(res_scDC_noClust, c("cond1","cond1","cond1","cond1",
                              "cond2","cond2","cond2","cond2"))
```

```{r}
densityCI(res_scDC_noClust, c("cond1","cond1","cond1","cond1",
                              "cond2","cond2","cond2","cond2"))
```

```{r}
source("glm.R")
res_GLM <- fitGLM_fixed(res_scDC_noClust, c("cond1","cond1","cond1","cond1",
                                      "cond2","cond2","cond2","cond2"), 
                  pairwise = FALSE)
```

with clustering

```{r}
res_scDC_clust = scDC_clustering(exprsMat, cellTypes, 
                                   subject, calCI = TRUE, 
                                   calCI_method = c("percentile", "BCa", "multinom"))
```

